\documentclass{article}
\usepackage{amsmath, amssymb, xcolor, ulem, tikz-cd}
\newcommand{\n}{\\ \\}
\renewcommand{\sp}{\phantom{a}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\cost}{\text{cost}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\total}{\text{total}}
\renewcommand{\lg}{\log_2}
\newcommand{\blue}[1]{{\color{blue}{#1}}}
\begin{document}
send \sout{memes} comments to tomachello3@gmail.com\n
\textbf{amortised cost}
example. you're given an $n$ bit register, and the operation $inc:x\mapsto x+1$. it's reasonable to say $\cost(inc,x)=\#\{\text{lsb bits} =1\}$. so at worst $\cost=n$. on the other hand, $\E\cost=\sum k/2^k$ is small. but also note the following: if you had $\cost = k$ then it will take a exponential time for $\cost$ to equal $k$ again. \blue{one costly operation for a bunch of cheap ones.}
\n\n
\textbf{Tarjan's potential method}
let  $\phi:S\mapsto\Z_{\ge0}$ assign structures a non-negative integer. if $a=t+\Delta\phi$ where $t$ is the time it takes to make the change $\Delta S$. then $\sum a + \phi(S_0)\ge t_{\total}$.\\
above example contd. let $\phi=\total\sp \#1$'s in register. if $\cost=k$ then $t=k$ but $\Delta\phi=1-k$ (or $-k$ if $k=n$) so that $a=1$ and $t_\total\le\#\text{operations}+n$.
\n\n
\textbf{credit method}
give each item some dollars. if an operation costs $t$, it has to be payed for by the items. if we can make this work, $t_\total\le$starting total money.
\n\n
\textbf{binomial heap}
$B_0=\bullet$ and $B_{i+1}=(1+r)B_i=\begin{tikzcd}
	& \bullet \\
	& {B_i} &&& \bullet \\
	\bullet && \bullet && {B_i} \\
	&&& \bullet && \bullet
	\arrow[no head, from=1-2, to=3-1]
	\arrow[no head, from=1-2, to=3-3]
	\arrow[no head, from=3-1, to=3-3]
	\arrow[no head, from=1-2, to=2-5]
	\arrow[no head, from=2-5, to=4-6]
	\arrow[no head, from=2-5, to=4-4]
	\arrow[no head, from=4-6, to=4-4]
\end{tikzcd}$ in $B_n$ we have $2^n$ vertices and height $n$. the root has $n$ edges and there's ${n\choose k}$ nodes of depth $k$. say we have $N=\sum a_j2^j$ keys. we keep a $B_j$ if $a_j=1$. it costs $1$ to make two $B_i$'s into one $B_{i+1}$ so it costs $\lg N$ to meld two heaps of size at most $N$ or insert a new key. it costs $\lg N$ to find min, and also to delete it (how?).
\clearpage
algo2 notes https://www.cs.technion.ac.il/~hamilis/Algorithms2.pdf\\
data structures 2 notes https://www.cs.technion.ac.il/~itai/Courses/ds2/lectures/lecture.html\\
\end{document}